# Udacity Data Analyst Nanodegree
# P05_Explonatory Data Analysis
  Martin Rypar
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(GGally)
library(RColorBrewer)

```

```{r echo=FALSE, Load_the_Data}

setwd("C:/Users/mrecl/Desktop/Exploratory_Data_Analysis/project/P5_Explonatory_Data_Analysis")
data <- read.csv("P5_UCI_Credit_Card_Dataset.csv", sep = ",")

```

# UCI Credit Card Dataset

> This dataset contains information on default payments, demographic factors, 
credit data, history of payment, and bill statements of credit card clients in 
Taiwan from April 2005 to September 2005. The dataset was downloaded from https://www.kaggle.com/uciml/default-of-credit-card-clients-dataset and further
informations are avaliable in enclosed P05_Dataset_Description.txt file.

# Preparation and getting some basic intuition of dataset

```{r echo=FALSE, message=FALSE, warning=FALSE, Intuition}
# convert gender
for (i in 1:length(data$SEX))
  if (data$SEX[i] == 1){
    data$SEX[i] = "male"
  } else if (data$SEX[i] == 2){
    data$SEX[i] = "female"
  } else
    data$SEX[i] = NULL

# convert education
for (i in 1:length(data$EDUCATION))
  if (data$EDUCATION[i] == 1){
    data$EDUCATION[i] = "graduate school"
  } else if (data$EDUCATION[i] == 2){
    data$EDUCATION[i] = "university"
  } else if (data$EDUCATION[i] == 3){
    data$EDUCATION[i] = "high school"
  } else {
    data$EDUCATION[i] = "others"
  }

# convert marriage
for (i in 1:length(data$MARRIAGE))
  if (data$MARRIAGE[i] == 1){
    data$MARRIAGE[i] = "married"
  } else if (data$MARRIAGE[i] == 2){
    data$MARRIAGE[i] = "single"
  } else if (data$MARRIAGE[i] == 3){
    data$MARRIAGE[i] = "divorced"
  } else{
    data$MARRIAGE[i] = "others"
  }

names(data)[25] = "default"

data$default_b <- data$default
# convert default
for (i in 1:length(data$default))
  if (data$default[i] == 1){
    data$default[i] = "yes"
  } else if (data$default[i] == 0){
    data$default[i] = "no"
  } else
    data$default[i] = NULL

# convert PAY_0 - PAY_6
names = names(data)[7:12]
for (i in c(7:12))
  for (j in 1:length(data$PAY_0))
    if (data[[i]][[j]] == -2){
      data[[i]][[j]] = "no consumption"
    } else if (data[[i]][[j]] == -1){
      data[[i]][[j]] = "paid"
    } else if (data[[i]][[j]] == 0){
      data[[i]][[j]] = "revolving"
    } else if (data[[i]][[j]] == 1){
      data[[i]][[j]] = "1m"
    } else if (data[[i]][[j]] == 2){
      data[[i]][[j]] = "2m"
    } else if (data[[i]][[j]] == 3){
      data[[i]][[j]] = "3m"
    } else if (data[[i]][[j]] == 4){
      data[[i]][[j]] = "4m"
    } else if (data[[i]][[j]] == 5){
      data[[i]][[j]] = "5m"
    } else if (data[[i]][[j]] == 6){
      data[[i]][[j]] = "6m"
    } else if (data[[i]][[j]] == 7){
      data[[i]][[j]] = "7m"
    } else if (data[[i]][[j]] == 8){
      data[[i]][[j]] = "8m"
    } else{
      data[[i]][[j]] = "NA"
    }

# weighted averages as a new variables
data$BILL_AMT <- (6*data$BILL_AMT1 + 5*data$BILL_AMT2 + 4*data$BILL_AMT3 + 3*data$BILL_AMT4 + 2*data$BILL_AMT5 + data$BILL_AMT6)/(21)
data$PAY_AMT <- (6*data$PAY_AMT1 + 5*data$PAY_AMT2 + 4*data$PAY_AMT3 + 3*data$PAY_AMT4 + 2*data$PAY_AMT5 + data$PAY_AMT6)/(21)

str(data)
summary(data)

```

# Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}

ggplot(aes(default), data = data) +
  stat_count(aes(fill = default)) +
  scale_y_continuous(breaks = seq(0, 30000, 2000))
     
ggplot(aes(SEX), data = data) +
  stat_count(aes(fill = SEX))

ggplot(aes(EDUCATION), data = data) + 
  stat_count(aes(fill = EDUCATION))

ggplot(aes(MARRIAGE), data = data) + 
  stat_count(aes(fill = MARRIAGE))

ggplot(aes(PAY_0), data = data) +
  geom_bar(stat = "count") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 15000, 1000)) +
  xlab("Repayment status")

# QUANTITATIVE VARIABLES

ggplot(aes(AGE), data = data) + 
  geom_density(aes(color = default, fill = default), stat = "bin", binwidth = 5) +
  scale_x_continuous(breaks = seq(20, 80, 5)) 

ggplot(aes(AGE), data = data) + 
  geom_density(aes(color = default, fill = default), stat = "bin", binwidth = 5) +
  scale_y_log10()
  
ggplot(aes(AGE), data = data) +
  geom_density(aes(color = default, fill = default), position = "fill", stat = "bin", binwidth = 5) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  scale_x_continuous(breaks = seq(20, 80, 5)) +
  geom_hline(yintercept = c(0.7, 0.8), linetype = 5)

ggplot(aes(LIMIT_BAL), data = data) +
  geom_histogram(binwidth = 25000) +
  scale_x_continuous(breaks = seq(0, 1000000, 50000), limits = c(0, quantile(data$LIMIT_BAL, 0.99)))

ggplot(aes(LIMIT_BAL), data = data) +
  geom_histogram() +
  scale_x_log10()

ggplot(aes(BILL_AMT), data = data) +
  geom_histogram(binwidth = 25000) +
  scale_x_continuous(breaks = seq(0, 500000, 50000), limits = c(0, quantile(data$BILL_AMT, 0.99)))

ggplot(aes(BILL_AMT), data = data) +
  geom_histogram() +
  scale_x_log10()

ggplot(aes(PAY_AMT), data = data) +
  geom_freqpoly(bins = 10) +
  scale_x_continuous(breaks = seq(0, 100000, 5000), limits = c(0, quantile(data$PAY_AMT, 0.99)))

ggplot(aes(PAY_AMT), data = data) +
  geom_freqpoly() +
  scale_x_log10()

```

# Univariate Analysis

> The dataset consists of 30 000 observations and 25 variables (age, sex,
education, marriage, amount of given credit, repayment status 1-6 months back, 
amount of bill statement 1-6 months back, 
amount of previous payments 1-6 months back, default) where I am mainly
interested in how is default variable (customer will not be able to pay 
his/her debts in next month) affected by other variables.
Closer description can be found in enclosed P05_Description_Dataset.txt file.

> First I recoded some of the categorical variables 
from their previous numerical 
coding (dummy variables) to more explicit and easy to read naming in order to
have more clear plot descriptions.

> Next I created two new variables:
From amount of bill statements for last 6 months (BILL_AMT) were calculated 
weighted averages for each customer and similarly for amount of previous 
payments (PAY_AMT) for last 6 months.

### Categorical Variables
* ~22% is general proportion of defaulted customers within dataset
* More than a half of the dataset consists of females  
* The most usual customer has university degree 
* More than a half of customers are single
* The majority of people use revolving credit to pay debts

### Quantitative Variables
* Age as skewed distribution (long-tail) with majority of customers
within 25-40 years old (older people do not use the credit cards and they prefer
cash in most of the cases), median age 34 and proportion of defaulted customers
remains between 0.2 and 0.3 with an exception of customers older than 70 years,
where is huge gap and approx. only a half of people around 75 y.o. paid 
their debts.
* LIMIT_BAL (amount of given credit) variable seems to have roughly 
skewed distribution
* Average amount of bill statement (BILL_AMT) has the skewed distribution
* The most common average of previous payments (PAY_AMT) is ~ 2000 NT dollars
and again is of skewed distribution


# Bivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}

ggplot(aes(SEX), data = data) +
  geom_bar(aes(fill = default), position = "fill", stat = "count") +
  scale_y_continuous(breaks = seq(0,1,0.1))

# as others and unknown are very small samples and thus there is no relevant signal from them I decided to remove them
ggplot(aes(EDUCATION), data = subset(data, EDUCATION != "others")) +
  geom_bar(aes(fill = default), position = "fill", stat = "count") +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

# as others is very small sample and thus there is no relevant signal from it I decided to remove it
ggplot(aes(MARRIAGE), data = subset(data, MARRIAGE != "others")) +
  geom_bar(aes(fill = default), position = "fill", stat = "count") +
  scale_y_continuous(breaks = seq(0,1,0.1))

ggplot(aes(PAY_0), data = data) +
  geom_bar(aes(color = default, fill = default), stat = "count", position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

ggplot(aes(PAY_2), data = data) +
  geom_bar(aes(color = default, fill = default), stat = "count", position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

ggplot(aes(PAY_3), data = data) +
  geom_bar(aes(color = default, fill = default), stat = "count", position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

ggplot(aes(default, LIMIT_BAL), data = data) +
  geom_boxplot(aes(color = default)) +
  # ylim(0, quantile(data$LIMIT_BAL, 0.95)) +
  scale_y_continuous(breaks = seq(0, 1000000, 50000))

ggplot(aes(default, BILL_AMT), data = data) +
  geom_boxplot(aes(color = default)) +
  ylim(0, quantile(data$BILL_AMT, 0.95))
  
ggplot(aes(default, PAY_AMT), data = data) +
  geom_boxplot(aes(color = default)) +
  ylim(0, quantile(data$PAY_AMT, 0.95))

ggplot(aes(SEX, PAY_AMT), data = data) +
  geom_boxplot(aes(color = SEX)) +
  scale_y_continuous(limits = c(0, quantile(data$PAY_AMT, 0.95)), breaks = seq(0, 50000, 5000))

ggplot(aes(PAY_0), data = data) +
  geom_bar(aes(fill = SEX), position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m"))

ggplot(aes(PAY_2), data = data) +
  geom_bar(aes(fill = SEX), position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m"))

ggplot(aes(PAY_3), data = data) +
  geom_bar(aes(fill = SEX), position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m"))

ggplot(aes(EDUCATION), data = subset(data, EDUCATION != "others")) +
  geom_bar(aes(fill = PAY_0), position = "fill") +
  scale_fill_brewer(palette = "Spectral")

ggplot(aes(MARRIAGE), data = subset(data, MARRIAGE != "others")) +
  geom_bar(aes(fill = PAY_0), position = "fill") +
  scale_fill_brewer(palette = "Spectral") +
  facet_wrap(~ EDUCATION)

# theme_set(theme_minimal(20))
# set.seed(1836)
# data.subset <- data[, c(2:8, 25:27)]
# names(data.subset)
# ggpairs(data.subset[sample.int(nrow(data.subset), 1000), ])

```

# Bivariate Analysis

* The proportion of
defaulted customers per sex is slightly bigger within males
* The proportion of defaulted 
customers per education seems to be highest within customers who completed only 
high school (but universtity graduates close enough)
* The biggest proportion of defaults showed divorced customers
* The most significant change in proportion of defaulted and non defaulted
customers repeatedly occured between 1 month delay in payment and 2 months delay
in payment - which draggs my attention
* LIMIT_BAL (amount of given credit) variable has higher median within non
defaulted customers
* Average amount of bill statement (BILL_AMT) do not shows any significant
differencies within people who default and those who do not
* As expected average amount of previous payments (PAY_AMT) has higher median
within people who do not default but with a higher variance
* Females seem to be better debtors as they pay on time more often then males
and the delay seems to be shorter as well for women


# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}

data$PAY <- paste(data$PAY_6, data$PAY_5, data$PAY_4, data$PAY_3, data$PAY_2, data$PAY_0, sep = ",")

ggplot(aes(x = AGE, y = LIMIT_BAL, color = default), data = data) + 
  geom_point()

ggplot(aes(AGE, PAY_AMT, color = default), data = data) +
  geom_point() +
  scale_y_continuous(limits = c(0, quantile(data$PAY_AMT, 0.95)))

# ggplot(aes(x = default, y = LIMIT_BAL, color = PAY_0), data = data) +
#   geom_boxplot() +
#   scale_fill_brewer(palette = "Spectral") +
#   scale_y_continuous(limits = c(0, quantile(data$LIMIT_BAL, 0.95)))
# 
# ggplot(aes(EDUCATION, fill = default), data = data) +
#   stat_count() +
#   facet_wrap(~ MARRIAGE)


# Logistic Regression Model

train_data <- data[c(2:7, 26:28)][1:24000,]
test_data <- data[c(2:7, 26:28)][24001:30000,]


model <- glm(default_b ~., family = binomial(link="logit"), data = train_data)

summary(model)

anova(model, test = "Chisq")

# library(pscl)
# pR2(model)

fitted.results <- predict(model,newdata=test_data, type='response')
fitted.results <- ifelse(fitted.results > 0.45,1,0)

misClasificError <- mean(fitted.results != test_data$default)
print(paste('Accuracy',1-misClasificError))

library(ROCR)
p <- predict(model, test_data, type="response")
pr <- prediction(p, test_data$default)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc <- round(auc, 4)
print (paste("Accuracy: ", auc))

```

# Multivariate Analysis

* Observation that customers with higher initial deposit default in less cases 
than customers with lower deposit was strenghten

* No clear evidence of that previous payments are dependent on age of customers

> As the decision if the customer will default in next month 
is classification task
with two possible outputs I tried to model this task
with a logistic regression model
(binary logistic regression).
First I very simply divide dataset into two groups: 
training data - on which I will train the model and estimate its parameters and
test data - on which I will test the accuracy of its predictions.

> The majority of features are statistical significant 
at least at the 0.05 level. But some of the features are not,
such as PAY_5m, PAY_6m, 
PAY_8m EDUCATIONhigh school, EDUCATIONuniversity, MARRIAGEmarried, 
MARRIAGEsingle, AGE.

> According to absolute values of estimates of categorical features 
it seems that PAY_0 has strong impact on the output of model. 
The SEX feature increases value of function by 0.1486 for males which means
that within males is slightly bigger chance of default, if the SEX is female
then value of function do not change (this relationship is partially reflected
in intercept).
The SEX features is in conformity with my previous findings, on the other hand
within the EDUCATION features it does not seem so. My findigs are that 
high school customers are more likely to default opposite to 
university customers whereas in the model 
it is contrarily. That might points out on that the model is not perfect.

> Model confirm my findigns from previous
parts such as: the higher initial deposit (LIMIT_BAL) the lower probability of 
default or males are more likely to default than females. 
The difference between the null deviance and the residual deviance shows how the
model is doing against the null model (a model with only the intercept). 
The wider this gap, the better. Noticeable decrease of Residual Deviance caused
adding PAY_0 (repayment status) feature into model 
and also LIMIT_BAL (initial deposit) feature.

> Accuracy 0.8295 represents how precise is model when predicting default status 
on a new set of data.
AUC (area under the curve) is another measure of accuracy, typical for 
classification tasks ranging from 0 to 1 (closer to 1 is better model).
The ROC is a curve generated by plotting the true positive rate (TPR) against 
the false positive rate (FPR) at various threshold settings 
while the AUC is the area under the ROC curve.
My model shows AUC 0.76 which is better than random guessing (AUC 0.5)
but in general 0.76 is not very precise and  at this kind of task is demanded
much better precision thus the model could be significantly improved.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

ggplot(aes(default, LIMIT_BAL), data = data) +
  geom_boxplot(aes(color = default)) +
  ggtitle("Initial deposit per default status") +
  scale_y_continuous(breaks = seq(0, 1000000, 50000))

```

### Description One
> The first plot shows how initial deposit(LIMIT_BAL) vary between
default statuses, even though it is simple plot it shows valuable information
that non defaulted customers have higher median of the deposit compared to 
customers who defaulted. And this information could be very quickly used by bank
or other financial institution exposed to credit risk to for example analyze
minimal level of initial deposit.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

ggplot(aes(PAY_0), data = data) +
  geom_bar(aes(fill = SEX), position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  ggtitle("Proportion of customers per repayment status and sex")  

```

### Description Two
> I like this one because it nicely develops the finding
that females are better 
in repaying their debts within this dataset. Where the proportion in groups
like no consumption, revolving and paid are clearly bigger within females.
Also trend seems to prompt that males have longer delays in repayment. 
This was observed also in different months (PAY_2, PAY_3, ...).

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

ggplot(aes(PAY_0), data = data) +
  geom_bar(aes(color = default, fill = default), stat = "count", position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m", "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  ggtitle("Proportion of customers per default and repayment status")

```

### Description Three
> The third plot is very similar to second one, 
but brings very different insights
and also offers one interesting thing. The plot shows proportion of defaulted
customers per repayment status in last month, where is clear that customers with
longer delay will default with much bigger probability. The interesting 
thing is that not only in the last month but even in previous months there is
a big gap in proportion of defaults between one month dealay and 
two months delay. It is sign that if customer is not able to repay his/her debt 
until one month delay, the chances of repayment are dramatically dragged down.

------

# Reflection

> A lot of insights are very intuitive and could be expected, nevertheless
thats the one of the reasons of doing such analysis - to confirm or disprove  
the ideas, intuitions and expectations about the data. 
After few projects in Python a was quite comfortable with the language and 
thus the change of the environment was a bit painfull for me even though 
I like R and I like how easy is to do the nice plots in it. 
And also during analysis I struggled a little bit with continuity 
as I often leave the
analysis and then return to it after few days and 
I felt like I lost the connection 
and I had to remind myself all the previous parts. On the other hand it has 
also positive side
that it could bring some new points of view and ideas.

> In future work I would like to develop better model 
by tuning the one I used -
as my approach was quite simple (e.g. division of dataset could be done better
as accuracy is dependent on the manual split to training and test set,
prepare better features)
or develop more advanced and more sophisticated
classification models (neural networks, naive bayes or decision trees). 


# Sources

http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
https://www.kaggle.com/moscolitos/exploratory-data-analysis/code
https://s3.amazonaws.com/content.udacity-data.com/courses/ud651/diamondsExample_2016-05.html
https://stackoverflow.com/questions/31561238/lapply-function-loops-on-list-of-lists-r
http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
https://www.r-bloggers.com/how-to-perform-a-logistic-regression-in-r/
http://r-statistics.co/Logistic-Regression-With-R.html
