# Udacity Data Analyst Nanodegree
# P5_Explonatory Data Analysis
 Martin Rypar
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(GGally)
library(RColorBrewer)

```

```{r echo=FALSE, Load_the_Data}

setwd("C:/Users/mrecl/Desktop/Exploratory_Data_Analysis/project")
setwd("P5_Explonatory_Data_Analysis")
data <- read.csv("P5_UCI_Credit_Card_Dataset.csv", sep = ",")

```

## UCI Credit Card Dataset

This dataset contains information on default payments, demographic factors, 
credit data, history of payment, and bill statements of credit card clients in 
Taiwan from April 2005 to September 2005. The dataset was downloaded from https://www.kaggle.com/uciml/default-of-credit-card-clients-dataset and further
informations are avaliable in enclosed P05_Dataset_Description.txt file.

## Preparation and getting some basic intuition of dataset

```{r echo=FALSE, message=FALSE, warning=FALSE, Intuition}
# convert gender
for (i in 1:length(data$SEX))
  if (data$SEX[i] == 1){
    data$SEX[i] = "male"
  } else if (data$SEX[i] == 2){
    data$SEX[i] = "female"
  } else
    data$SEX[i] = NULL

# convert education
for (i in 1:length(data$EDUCATION))
  if (data$EDUCATION[i] == 1){
    data$EDUCATION[i] = "graduate school"
  } else if (data$EDUCATION[i] == 2){
    data$EDUCATION[i] = "university"
  } else if (data$EDUCATION[i] == 3){
    data$EDUCATION[i] = "high school"
  } else {
    data$EDUCATION[i] = "others"
  }

# convert marriage
for (i in 1:length(data$MARRIAGE))
  if (data$MARRIAGE[i] == 1){
    data$MARRIAGE[i] = "married"
  } else if (data$MARRIAGE[i] == 2){
    data$MARRIAGE[i] = "single"
  } else if (data$MARRIAGE[i] == 3){
    data$MARRIAGE[i] = "divorced"
  } else{
    data$MARRIAGE[i] = "others"
  }

names(data)[25] = "default"

data$default_b <- data$default
# convert default
for (i in 1:length(data$default))
  if (data$default[i] == 1){
    data$default[i] = "yes"
  } else if (data$default[i] == 0){
    data$default[i] = "no"
  } else
    data$default[i] = NULL

# convert PAY_0 - PAY_6
names = names(data)[7:12]
for (i in c(7:12))
  for (j in 1:length(data$PAY_0))
    if (data[[i]][[j]] == -2){
      data[[i]][[j]] = "no consumption"
    } else if (data[[i]][[j]] == -1){
      data[[i]][[j]] = "paid"
    } else if (data[[i]][[j]] == 0){
      data[[i]][[j]] = "revolving"
    } else if (data[[i]][[j]] == 1){
      data[[i]][[j]] = "1m"
    } else if (data[[i]][[j]] == 2){
      data[[i]][[j]] = "2m"
    } else if (data[[i]][[j]] == 3){
      data[[i]][[j]] = "3m"
    } else if (data[[i]][[j]] == 4){
      data[[i]][[j]] = "4m"
    } else if (data[[i]][[j]] == 5){
      data[[i]][[j]] = "5m"
    } else if (data[[i]][[j]] == 6){
      data[[i]][[j]] = "6m"
    } else if (data[[i]][[j]] == 7){
      data[[i]][[j]] = "7m"
    } else if (data[[i]][[j]] == 8){
      data[[i]][[j]] = "8m"
    } else{
      data[[i]][[j]] = "NA"
    }

# weighted averages as a new variables
data$BILL_AMT <- (6*data$BILL_AMT1 + 5*data$BILL_AMT2 + 4*data$BILL_AMT3 +
                    3*data$BILL_AMT4 + 2*data$BILL_AMT5 + data$BILL_AMT6)/(21)
data$PAY_AMT <- (6*data$PAY_AMT1 + 5*data$PAY_AMT2 + 4*data$PAY_AMT3 +
                   3*data$PAY_AMT4 + 2*data$PAY_AMT5 + data$PAY_AMT6)/(21)

str(data)
summary(data)

```

## Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_1}

ggplot(aes(default), data = data) +
  stat_count() +
  scale_y_continuous(breaks = seq(0, 30000, 2000))
 
```

As I am mainly interested in default variable, lets look at number of default
vs non default customers.
It is clear that default variable is not equally distributed, as
most of the customers will not default.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_2}
    
ggplot(aes(SEX), data = data) +
  stat_count()

```

Now lets look at the categorical variables.
One would expect similar number of males and females within dataset,
but here is a quite gap between those two on behalf of females.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_3}

ggplot(aes(EDUCATION), data = data) + 
  stat_count()

```

Education features is neither equally distributed as majority of people
attended university (university, graduate school) and others variable which
has arguable information value but I decided to keep those observations in.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_4}

ggplot(aes(MARRIAGE), data = data) + 
  stat_count()

```

The majority of dataset is consisted of similar number of single 
and married customers, only a fraction of dataset are divorced and again 
others which meaning is not completely clear again.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_5}

ggplot(aes(PAY_0), data = data) +
  geom_bar(stat = "count") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 15000, 1000)) +
  xlab("Repayment status")

```

Possibly the most important feature - repayment status.
Plot showing repayment status of customer in last recorded month (September),
(-2: No consumption; 
-1: Paid in full; 
0:The use of revolving credit; 
1 = payment delay for one month; 
2 = payment delay for two months; . . .; 
8 = payment delay for eight months; 
9 = payment delay for nine months and above)
Distribution almost like a skewed, if we would not take into account
no consumption. Plot also suggests that credit card users pay their 
debts most often by using revolving credit. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_6}

ggplot(aes(AGE), data = data) + 
  geom_density(stat = "bin", binwidth = 5) +
  scale_x_continuous(breaks = seq(20, 80, 5))

```

And now quantitative features.
Clearly skewed distribution of age features shows that majority of people
using credit cards are people ~ 25-40 years old. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_7}

ggplot(aes(LIMIT_BAL), data = data) +
  geom_histogram(binwidth = 25000) +
  scale_x_continuous(breaks = seq(0, 1000000, 50000),
                     limits = c(0, quantile(data$LIMIT_BAL, 0.99)))

```

LIMIT_BAL feature shows histogram of initial deposits where the distribution
might remind skewed distribution. This variable may be important as
people with bigger initial deposit might be less prone to default.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_8}

ggplot(aes(LIMIT_BAL), data = data) +
  geom_histogram() +
  scale_x_log10()

```

Same plot as before only with after log transformation of x axis. 
Not particularly helpful.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_9}

ggplot(aes(BILL_AMT), data = data) +
  geom_histogram(binwidth = 25000) +
  scale_x_continuous(breaks = seq(0, 500000, 50000),
                     limits = c(0, quantile(data$BILL_AMT, 0.99)))

```

Plot depicts histogram of different averages of amounts of bill statement, 
where it shows again skewed distribution. I am interested in this feature 
as it could cluster customers by their amount of bill statement and thus might 
contain signal whether is some cluster more prone to default.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_10}

ggplot(aes(PAY_AMT), data = data) +
  geom_freqpoly(bins = 10) +
  scale_x_continuous(breaks = seq(0, 100000, 5000),
                     limits = c(0, quantile(data$PAY_AMT, 0.99)))

```

Histogram with average volume of previous payments 
again with strongly skewed distribution,
in conformity with plot 5, 
which showed that most of the people use
revolving credit to repay 
and thus do not pay directly which is depicted here.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_11}

ggplot(aes(PAY_AMT), data = data) +
  geom_freqpoly() +
  scale_x_log10()

```

After log transformation of x axis of the previous plot, we can see
plot more similar to normal distribution.


## Univariate Analysis

The dataset consists of 30 000 observations and 25 variables (age, sex,
education, marriage, amount of given credit, repayment status 1-6 months back, 
amount of bill statement 1-6 months back, 
amount of previous payments 1-6 months back, default) where I am mainly
interested in how is default variable (customer will not be able to pay 
his/her debts in next month) affected by other variables.
Closer description can be found in enclosed P05_Description_Dataset.txt file.

First I recoded some of the categorical variables 
from their previous numerical 
coding (dummy variables) to more explicit and easy to read naming in order to
have more clear plot descriptions.

Next I created two new variables:
From amount of bill statements for last 6 months (BILL_AMT) were calculated 
weighted averages for each customer with biggest weight on the most recent 
record and smallest weight on the oldest record. 
Similarly were calculated weighted average PAY_AMT of 
payments for last 6 months.

#### Categorical Variables
* ~22% is general proportion of defaulted customers within dataset
* More than a half of the dataset consists of females  
* The most usual customer has university degree 
* More than a half of customers are single
* The majority of people use revolving credit to pay debts

#### Quantitative Variables
* Age as skewed distribution (long-tail) with majority of customers
within 25-40 years old (older people do not use the credit cards and they prefer
cash in most of the cases), median age 34 and proportion of defaulted customers
remains between 0.2 and 0.3 with an exception of customers older than 70 years,
where is huge gap and approx. only a half of people around 75 y.o. paid 
their debts.
* LIMIT_BAL (amount of given credit) variable seems to have roughly 
skewed distribution
* Average amount of bill statement (BILL_AMT) has the skewed distribution
* The most common average of previous payments (PAY_AMT) is ~ 2000 NT dollars
and again is of skewed distribution

In general the goal of ploting all the plots were to get some intuition 
and basic understnading of how the dataset looks like and find out 
what are distributions of particular features.


## Bivariate Plots Section

As I have mentioned before, I am mainly interested wheter customer will default
or not in the following month.
So now I am going to look at the differences within customers 
divided per certain characteristics and 
wheter they default or not, in other words
I am going to look at the proportions of defaulted vs non defaulted based on
their attributes such as age, sex, education etc.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots1}

ggplot(aes(SEX), data = data) +
  geom_bar(aes(fill = default), position = "fill", stat = "count") +
  scale_y_continuous(breaks = seq(0,1,0.1))

```

One can notice that within females is default rate slightly lower ~0.21
than within males ~0.24, which might indicate that females are more 
responsible when repaying debts.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots2}
# as others and unknown are very small samples 
# and thus there is no relevant signal from them I decided to remove them
ggplot(aes(EDUCATION), data = subset(data, EDUCATION != "others")) +
  geom_bar(aes(fill = default), position = "fill", stat = "count") +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

```

Also education could be one of the important features helping to estimate
wheter customer will default or not. Hypothesis is that customers with a better
education tend to default in lower proportion than customers with a worse 
education.
Some differencies might be noticed
when looking at the plot, where customers finished 
graduate school (the highest academical degree within dataset) have 
the lowest proportion of defaults ~0.19 and people finished 
only high school (the lowest academical degree within dataset) have the highest
proportion of defaults ~0.25. University absolvents somewhere between ~0.23 and
others were excluded from plot due to lack of records.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots3}
# as others is very small sample
# and thus there is no relevant signal from it I decided to remove it
ggplot(aes(MARRIAGE), data = subset(data, MARRIAGE != "others")) +
  geom_bar(aes(fill = default), position = "fill", stat = "count") +
  scale_y_continuous(breaks = seq(0,1,0.1))

```

Marriage variable could also help to estimate potential default of customers,
thats why I ploted this graph. Single customers with the lowest proportion of 
defaults ~0.21, married slightly worse ~0.23 and divorced ~0.26. Others were
again excluded - same reason as before.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots4}

ggplot(aes(AGE), data = data) + 
  geom_density(aes(color = default, fill = default),
               stat = "bin", binwidth = 1) +
  scale_x_continuous(breaks = seq(20, 80, 5)) 

```

Next I would like to see if age distribution is somehow different based on 
default and non default customers. Plot shows that the curves are similar, but
the default customers curve seems to decrease much slower, it has much smaller
slope opposite to non default customers. That might indicate that customer 
between 35 - 60 are more likely to default.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots5} 

ggplot(aes(AGE), data = data) +
  geom_density(aes(color = default, fill = default),
               position = "fill", stat = "bin", binwidth = 5) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  scale_x_continuous(breaks = seq(20, 80, 5)) +
  geom_hline(yintercept = c(0.7, 0.8), linetype = 5)

```

To examine my conjecture, lets look at more closely. There is no clear evidence
After a while playing with the binwidth, there is some signs of increasing 
proportion of defaults with increasing age within 35 - 60 range, 
but it is not clear evidence that older
people defaults more often than younger as the end 
of the curve is very volatile which could be caused by lack of observations.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots6}

ggplot(aes(PAY_0), data = data) +
  geom_bar(aes(color = default, fill = default),
           stat = "count", position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots7}

ggplot(aes(PAY_2), data = data) +
  geom_bar(aes(color = default, fill = default),
           stat = "count", position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots8}

ggplot(aes(PAY_3), data = data) +
  geom_bar(aes(color = default, fill = default),
           stat = "count", position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

```

To examine payment status for last 3 month and proportion of defaults I have
ploted the 3 above graphs, which showing that no within no consumption, 
revolving, paid and 1m delay is proportion of defaults clearly minimal ~0.2.
But with delay 2 months and more the proportion of defaults dramatically 
increased ~0.6. The most significant gap is between 1m delay and 2m delay,
which is sign that if customer is not able to repay the debt until 1m delay
then the chance of repayment rapidly decrease. 


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots9}

ggplot(aes(default, LIMIT_BAL), data = data) +
  geom_boxplot(aes(color = default)) +
  scale_y_continuous(breaks = seq(0, 1000000, 50000))

```

Do customers with different initial deposits diverse based on default
information? Yes, it seems so. 
Customers who defaulted have lower median of initial deposits opposite to them
who did not default.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots10}

ggplot(aes(default, BILL_AMT), data = data) +
  geom_boxplot(aes(color = default)) +
  ylim(0, quantile(data$BILL_AMT, 0.95))

```

But same which apply for initial deposit, does not apply for average
amount of bill statement. There median relatively same for both groups - default
and non default. Top 5% values were cut off to eliminate outliers.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots11}

ggplot(aes(default, PAY_AMT), data = data) +
  geom_boxplot(aes(color = default)) +
  ylim(0, quantile(data$PAY_AMT, 0.95))

```

And as could be expected the average amount of previous payments have higher 
median within non defaulted customers, which is quite obvious that customers
who pay their debts could not be in default. Top 5% values were cut off 
to eliminate outliers.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots12}

ggplot(aes(PAY_0), data = data) +
  geom_bar(aes(fill = SEX), position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m"))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots13}

ggplot(aes(PAY_2), data = data) +
  geom_bar(aes(fill = SEX), position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m"))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots14}

ggplot(aes(PAY_3), data = data) +
  geom_bar(aes(fill = SEX), position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m"))

```

To analyze not only default feature, but also other features I ploted those
3 graphs above, where is noticeable difference in males and females repayment 
morale. Females seem to be more responsible in repayments, the proportion of
females within no consumpion, revolving and paid is clearly bigger than males
proportion. Also females have shorter delays of repayments. Also it worths 
notice that there is a quite difference in count of females and males in the 
dataset on behalf of females.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots15}

ggplot(aes(EDUCATION), data = subset(data, EDUCATION != "others")) +
  geom_bar(aes(fill = PAY_0), position = "fill") +
  scale_fill_brewer(palette = "Spectral")

```

To get some intuition about repayment morale per education I generated
this plot above. Where clear dominance of revolving repayment is striking, only
within graduate school absolvents it looks like they are more conscientious
and instead of revolving they repaid the debt also significant part is
no consumption within this group and generaly it is obvious that best repayment
morale have group - gradute school. High school and university graduates shown
worse opposite to graduate school 
but with only minimal differencies between those two groups.


<!-- # theme_set(theme_minimal(20)) -->
<!-- # set.seed(1836) -->
<!-- # data.subset <- data[, c(2:8, 25:27)] -->
<!-- # names(data.subset) -->
<!-- # ggpairs(data.subset[sample.int(nrow(data.subset), 1000), ]) -->


## Bivariate Analysis

Within majority of features is proportion between default and non default
customers very small (sex, education, marriage) but within some of the features
we can find strong contrast (PAY_0, PAY_2, PAY_3) 
as those features could be much
stronger in discrimination of default and non default customers. 

I would expect that females are better in repaying their debts, but
I was slightly surprised by about how much are they better and 
about how much better is their repayment morale at least within this dataset.
The rest of the features behave in conformity with my expectations.

What it worths to say is that this kind of data are related to culture and part
of the world from where it were collected as in western countries there is 
probably different handling with money opposite to eastern countries etc.
Also this data represents mainly younger and middle-aged generation ~20-55, 
where dataset primary consists of people that age.

####Key findings:
* The proportion of
defaulted customers per sex is slightly bigger within males
* The proportion of defaulted 
customers per education seems to be highest within customers who completed only 
high school (but universtity graduates close enough)
* The biggest proportion of defaults showed divorced customers
* The most significant change in proportion of defaulted and non defaulted
customers repeatedly occured between 1 month delay in payment and 2 months delay
in payment - which draggs my attention
* LIMIT_BAL (amount of given credit) variable has higher median within non
defaulted customers
* Average amount of bill statement (BILL_AMT) do not shows any significant
differencies within people who default and those who do not
* As expected average amount of previous payments (PAY_AMT) has higher median
within people who do not default but with a higher variance
* Females seem to be better debtors as they pay on time more often then males
and the delay seems to be shorter as well for women


## Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots1}

data$PAY <- paste(data$PAY_6, data$PAY_5, data$PAY_4,
                  data$PAY_3, data$PAY_2, data$PAY_0, sep = ",")

ggplot(aes(x = AGE, y = LIMIT_BAL, color = default), data = data) + 
  geom_point(alpha = 0.5)

ggplot(aes(AGE, PAY_AMT, color = default), data = data) +
  geom_point(alpha = 0.5) +
  scale_y_continuous(limits = c(0, quantile(data$PAY_AMT, 0.95)))

```

My idea was that maybe older people tends to have much bigger initial deposit
as they are probably more conservative so I took a shot at it. Similarly I did
for amount of paid money. Neither one of the plots does show 
any clear pattern, even though on the age~initial deposit plot seems that
default customers concentrate very close to x-axis with minimal initial deposit.
Which might be interpreted as people with lower initial deposit are more likely
to default.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots2}

# ggplot(aes(x = default, y = LIMIT_BAL, color = PAY_0), data = data) +
#   geom_boxplot() +
#   scale_fill_brewer(palette = "Spectral") +
#   scale_y_continuous(limits = c(0, quantile(data$LIMIT_BAL, 0.95)))
# 
# ggplot(aes(EDUCATION, fill = default), data = data) +
#   stat_count() +
#   facet_wrap(~ MARRIAGE)
#
# ggplot(aes(EDUCATION), data = subset(data, EDUCATION != "others")) +
#   geom_bar(aes(fill = PAY_0), position = "fill") +
#   scale_fill_brewer(palette = "Spectral") +
#   facet_wrap(~ default)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots3}
# Logistic Regression Model

train_data <- data[c(2:7, 26:28)][1:24000,]
test_data <- data[c(2:7, 26:28)][24001:30000,]


model <- glm(default_b ~., family = binomial(link="logit"),
             data = train_data)

summary(model)

```

Before creating a model I split the data into train set and test set and 
afterwards used glm() function to create logistic regression model with
summary above.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots4}

anova(model, test = "Chisq")

```

Then to find out how particular features help to reduce residual deviance
in other words, how well they are helpful with improvement of the model, 
ANOVA summary is showed above.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots5}

# library(pscl)
# pR2(model)

fitted.results <- predict(model,newdata=test_data, type='response')
fitted.results <- ifelse(fitted.results > 0.45,1,0)

misClasificError <- mean(fitted.results != test_data$default)
print(paste('Accuracy',1-misClasificError))

```

After fitting model to test set to find out how well the model predicts
wheter customer will default or not, accuracy is calculated.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots6}

library(ROCR)
p <- predict(model, test_data, type="response")
pr <- prediction(p, test_data$default_b)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc <- round(auc, 4)
print (paste("AUC Accuracy: ", auc))

```

The ploted curve, called ROC, generated by plotting 
the true positive rate (TPR) against 
the false positive rate (FPR) at various threshold settings and 
AUC Accuracy are different measures of accuracy of the model, 
widely used for classification tasks.


## Multivariate Analysis

To sum up the first two plots:
Observation that customers with higher initial deposit default in less cases 
than customers with lower deposit was strenghten. 
No clear evidence of that previous payments are dependent on age of customers.

As the decision if the customer will default in next month 
is classification task
with two possible outputs I tried to model this task
with a logistic regression model
(binary logistic regression).
First I very simply divide dataset into two groups: 
training data - on which I will train the model and estimate its parameters and
test data - on which I will test the accuracy of its predictions.

The majority of features are statistical significant 
at least at the 0.05 level. But some of the features are not,
such as PAY_5m, PAY_6m, 
PAY_8m EDUCATIONhigh school, EDUCATIONuniversity, MARRIAGEmarried, 
MARRIAGEsingle, AGE.

According to absolute values of estimates of categorical features 
it seems that PAY_0 has strong impact on the output of model. 
The SEX feature increases value of function by 0.1486 for males which means
that within males is slightly bigger chance of default, if the SEX is female
then value of function do not change (this relationship is partially reflected
in intercept).
The SEX features is in conformity with my previous findings, on the other hand
within the EDUCATION features it does not seem so. My findigs are that 
high school customers are more likely to default opposite to 
university customers whereas in the model 
it is contrarily. That might points out on that the model is not perfect.

Model confirm my findigns from previous
parts such as: the higher initial deposit (LIMIT_BAL) the lower probability of 
default or males are more likely to default than females. 
The difference between the null deviance and the residual deviance shows how the
model is doing against the null model (a model with only the intercept). 
The wider this gap, the better. Noticeable decrease of Residual Deviance caused
adding PAY_0 (repayment status) feature into model 
and also LIMIT_BAL (initial deposit) feature.

Accuracy 0.8295 represents how precise is model when predicting default status 
on a new set of data.
AUC (area under the curve) is another measure of accuracy, typical for 
classification tasks ranging from 0 to 1 (closer to 1 is better model).
The ROC is a curve generated by plotting the true positive rate (TPR) against 
the false positive rate (FPR) at various threshold settings 
while the AUC is the area under the ROC curve.
My model shows AUC 0.76 which is better than random guessing (AUC 0.5)
but in general 0.76 is not very precise and  at this kind of task is demanded
much better precision thus the model could be significantly improved.

------

## Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

ggplot(aes(default, LIMIT_BAL), data = data) +
  geom_boxplot(aes(color = default)) +
  ggtitle("Initial deposit per default status") +
  scale_y_continuous(breaks = seq(0, 1000000, 50000))

```

### Description One
The first plot shows how initial deposit(LIMIT_BAL) vary between
default statuses, even though it is simple plot it shows valuable information
that non defaulted customers have higher median of the deposit compared to 
customers who defaulted. And this information could be very quickly used by bank
or other financial institution exposed to credit risk to for example analyze
minimal level of initial deposit.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

ggplot(aes(PAY_0), data = data) +
  geom_bar(aes(fill = SEX), position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  ggtitle("Proportion of customers per repayment status and sex")  

```

### Description Two
I like this one because it nicely develops the finding
that females are better 
in repaying their debts within this dataset. Where the proportion in groups
like no consumption, revolving and paid are clearly bigger within females.
Also trend seems to prompt that males have longer delays in repayment. 
This was observed also in different months (PAY_2, PAY_3, ...).

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

ggplot(aes(PAY_0), data = data) +
  geom_bar(aes(color = default, fill = default),
           stat = "count", position = "fill") +
  scale_x_discrete(limits = c("no consumption", "revolving", "paid", "1m",
                              "2m", "3m", "4m", "5m", "6m", "7m", "8m")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  ggtitle("Proportion of customers per default and repayment status")

```

### Description Three
The third plot is very similar to second one, 
but brings very different insights
and also offers one interesting thing. The plot shows proportion of defaulted
customers per repayment status in last month, where is clear that customers with
longer delay will default with much bigger probability. The interesting 
thing is that not only in the last month but even in previous months there is
a big gap in proportion of defaults between one month dealay and 
two months delay. It is sign that if customer is not able to repay his/her debt 
until one month delay, the chances of repayment are dramatically dragged down.

------

## Reflection

A lot of insights are very intuitive and could be expected, nevertheless
thats the one of the reasons of doing such analysis - to confirm or disprove  
the ideas, intuitions and expectations about the data. 
After few projects in Python a was quite comfortable with the language and 
thus the change of the environment was a bit painfull for me even though 
I like R and I like how easy is to do the nice plots in it. 
And also during analysis I struggled a little bit with continuity 
as I often leave the
analysis and then return to it after few days and 
I felt like I lost the connection 
and I had to remind myself all the previous parts. On the other hand it has 
also positive side
that it could bring some new points of view and ideas.

In future work I would like to develop better model 
by tuning the one I used -
as my approach was quite simple (e.g. division of dataset could be done better
as accuracy is dependent on the manual split to training and test set,
prepare better features)
or develop more advanced and more sophisticated
classification models (neural networks, naive bayes or decision trees). 


## Sources

http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
https://www.kaggle.com/moscolitos/exploratory-data-analysis/code
https://s3.amazonaws.com/content.udacity-data.com/
courses/ud651/diamondsExample_2016-05.html<br>
https://stackoverflow.com/questions/31561238/
lapply-function-loops-on-list-of-lists-r<br>
http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
https://www.r-bloggers.com/how-to-perform-a-logistic-regression-in-r/
http://r-statistics.co/Logistic-Regression-With-R.html
